---
title: "Fraud_Detection"
output: html_document
---

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggridges)
library(readr) #For reaading CSV Files
library(purrr)
library(corrplot)
library(caTools)
library(caret)
library(randomForest) #Used for Random Forest
library(ROSE)
library(rpart) #Used for Decision Tree
library(rpart.plot)
library(smotefamily) #Used for SMOTE
library(tidyverse) # metapackage with lots of helpful functions
library(pROC) # to use with AUC
library(RColorBrewer)#used for chart
library(e1071)
library(scales) #used for chart
```



```{r}
credit_card=read.csv("creditcard.csv")
```


```{r}
nrow(credit_card)
str(credit_card)
credit_card$Class = factor(credit_card$Class,levels=c(0,1))
summary(credit_card)
sum(is.na(credit_card))
```



```{r}
#Get distribution of different classes
table(credit_card$Class)

#Percentage
prop.table(table(credit_card$Class))

#Pie Chart of credit card transaction
labels = c("legit","fraud")
labels = paste(labels,round(100*prop.table(table(credit_card$Class)),2))
labels = paste0(labels,"%")
labels


pie(table(credit_card$Class) , labels  , col = c("orange","red"), main = " Pie Chart for transactions ")

```



```{r}

ggplot(credit_card,aes(x=Class))+
  geom_bar(position ="dodge",fill="blue")+
  scale_y_continuous()+
  scale_x_discrete()+
  ggtitle("Distribution of Calss Variable")
```


```{r}

ggplot(credit_card,aes(x=Class, y=Amount,group=Class))+
  geom_boxplot()+
  ggtitle("Amount by Class")
```



```{r}

data_h<- gather(credit_card,type,value,2:29) #Pivot_longer
head(data_h)

data_h <- data_h%>% select(Time,type,value,Amount,Class)
data_h = sample_n(data_h,4000)
nrow(data_h)

ggplot(data_h, aes(x=value,y = Amount, color=type))+
  geom_jitter()+
  facet_grid(.~Class)+
  ggtitle("Distribution of Normal and Fraud Class")

head(data_h)


ggplot(data=data_h) + geom_bar(mapping = aes(x=type,fill=Class)) + theme(axis.text.x = element_text(angle = 90,hjust=1,vjust=0.5))


```


## No Model Prediction

```{r}

class(credit_card$Class)
pred = rep.int(0,nrow(credit_card))

pred = factor(pred,levels=c(0,1))
confusionMatrix(data = pred , reference = credit_card$Class)

```



## Taking a part of dataset(40%)

```{r}

set.seed(1)
credit_card = credit_card %>% sample_frac(0.4)

class(credit_card$Class)

ggplot(data = credit_card,aes(x=V1,y=V2,col=Class))+geom_point()+theme_bw()+scale_color_manual(values = c('dodgerblue2','red'))


```


## Create train and test set

```{r}

set.seed(123)

data_sample = sample.split(credit_card$Class,SplitRatio = 0.80)
train_data = subset(credit_card,data_sample==TRUE)
test_data = subset(credit_card,data_sample==FALSE)
nrow(train_data)
nrow(test_data)


```



## Balaning Data

* Random Over Sampling

```{r}



table(train_data$Class)

n_legit = 90978   
new_frac_legit = 0.50
new_n_total = n_legit / new_frac_legit
new_n_total

over_sampling = ovun.sample(Class ~ . ,
                            data = train_data,
                            method = "over",
                            N = new_n_total,
                            seed=2019)

oversampled_credit = over_sampling$data

table(oversampled_credit$Class)

ggplot(data = oversampled_credit,aes(x=V1,y=V2,col=Class))+
  geom_point()+
  theme_bw()+
  scale_color_manual(values = c('dodgerblue2','red'))



ggplot(data = oversampled_credit,aes(x=V1,y=V2,col=Class))+
  geom_point(position = position_jitter(width=0.2))+
  theme_bw()+
  scale_color_manual(values = c('dodgerblue2','red'))



```


* Random Ubder Sampling

```{r}

table(train_data$Class)

n_fraud = 161
new_frac_fraud = 0.50
new_n_total = n_fraud / new_frac_fraud

under_sampling = ovun.sample(Class ~ . ,
                            data = train_data,
                            method = "under",
                            N = new_n_total,
                            seed=2019)
undersampled_credit = under_sampling$data
table(undersampled_credit$Class)

ggplot(data = undersampled_credit,aes(x=V1,y=V2,col=Class))+
  geom_point()+
  theme_bw()+
  scale_color_manual(values = c('dodgerblue2','red'))


```

* Both RUS , ROS

```{r}


n_new = nrow(train_data)
fraction_fraud_new = 0.50

sampling_reult = ovun.sample(Class ~ . ,
                             data = train_data,
                             method = "both",
                             N = n_new,
                             p = fraction_fraud_new,
                             seed=2019)

sampled_credit = sampling_reult$data
table(sampled_credit$Class)

ggplot(data = sampled_credit,aes(x=V1,y=V2,col=Class))+
  geom_point(position = position_jitter(width=0.2))+
  theme_bw()+
  scale_color_manual(values = c('dodgerblue2','red'))
```



## SMOTE Preprocessing - Synthetic Minority Over-sampling Technique


```{r}


#install.packages("smotefamily")

nrow(train_data)
table(train_data$Class)
n0 = 90978
n1= 161
r0 = 0.6

ntimes = ((1-r0) / r0) * (n0/n1) - 1 #432 times
#head(train_data)
#head(train_data[,-c(1,31)])
smote_output = SMOTE(X = train_data[,-c(1,31)],
                     target = train_data$Class,
                     K=5, #K is the number of nearest neighbours
                     dup_size = ntimes)

credit_smote = smote_output$data
table(credit_smote$Class)
nrow(credit_smote)
#changing class to Class

table(credit_smote$Class)
head(credit_smote)
names(credit_smote)[names(credit_smote) == "class"] <- "Class"
credit_smote$Class = as.factor(credit_smote$Class)
class(credit_smote$Class)

prop.table(table(credit_smote$Class))


```




```{r}

#Class distribution for original dataset

ggplot(train_data,aes(x=V1,y=V2,col=Class))+
  geom_point()+
  theme_bw()+
  scale_color_manual(values = c('dodgerblue2','red'))


ggplot(credit_smote,aes(x=V1,y=V2,col=Class))+
  geom_point()+
  theme_bw()+
  scale_color_manual(values = c('dodgerblue2','red'))
```


##### Decision Tree


```{r}

table(credit_smote$Class)

#Trees

head(credit_smote)

CART_model = rpart(Class ~ . ,credit_smote)
rpart.plot(CART_model,extra = 0,type=5,tweak=1.2)
predicted_val = predict(CART_model,test_data,type='class')
df = confusionMatrix(predicted_val,test_data$Class)
df$table
df
fourfoldplot(df$table)

precision <- df$byClass['Precision']    
precision
recall <- df$byClass['Recall']
recall
specificity<- df$byClass['Specificity']
specificity
F_Score = (2 * precision * recall) / (precision + recall)
paste("F-score",F_Score)
(F_Score)
paste(precision,recall,specificity)


```




#### Decision tree on original data

```{r}

CART_model = rpart(Class ~ . ,train_data[,-1])
rpart.plot(CART_model,extra = 0,type=5,tweak=1.2)

predicted_val = predict(CART_model,test_data[-1],type='class')
df = confusionMatrix(predicted_val,test_data$Class)

precision <- df$byClass['Precision']    
precision
recall <- df$byClass['Recall']
recall
specificity<- df$byClass['Specificity']
specificity
F_Score = (2 * precision * recall) / (precision + recall)
paste("F-Score",F_Score)

df

```


### Using our trained smote model on whole data(unbalanced)


```{r}


CART_model = rpart(Class ~ . ,credit_smote)

predicted_val = predict(CART_model,credit_card[-1],type='class')
df = confusionMatrix(predicted_val,credit_card$Class)
df
df$table

precision <- df$byClass['Precision']    
precision
recall <- df$byClass['Recall']
recall
specificity<- df$byClass['Specificity']
specificity
F_Score = (2 * precision * recall) / (precision + recall)
paste("F-Score",F_Score)

```



* Decision Tree without SMOTE

```{r}
#Using trained model and tesing it

CART_model = rpart(Class ~ . ,train_data[,-1])

predicted_val = predict(CART_model,credit_card[-1],type='class')
df = confusionMatrix(predicted_val,credit_card$Class)
nrow(credit_smote)
df

precision <- df$byClass['Precision']    
precision
recall <- df$byClass['Recall']
recall
specificity<- df$byClass['Specificity']
specificity
F_Score = (2 * precision * recall) / (precision + recall)
paste("F-Score",F_Score)

```





```{r}
# Random Forest with SMOTE
set.seed(1)
credit_smote = credit_smote %>% sample_frac(0.5)
nrow(credit_smote)
model_rfn <-  randomForest(Class~., data = credit_smote,
                           ntree = 1000, nodesize = 20)
#testing sample trained model on test data
predict_rfn <- predict(model_rfn, test_data, type = "response")
df = confusionMatrix(predict_rfn, test_data[["Class"]])
df

precision <- df$byClass['Precision']    
precision
recall <- df$byClass['Recall']
recall
specificity<- df$byClass['Specificity']
specificity
F_Score = (2 * precision * recall) / (precision + recall)
paste("F-Score",F_Score)

#importance(model_rfn

options(repr.plot.width=5, repr.plot.height=4)
varImpPlot(model_rfn,
          sort = T,
           n.var=10,
           main="Top 10 Most Important Variables")
```


### Checking for out of bag error

```{r}
# Random Forest with SMOTE
model_rfn200 <-  randomForest(Class~., data = credit_smote,
                           ntree = 200, nodesize = 20)
#testing sample trained model on test data
predict_rfn <- predict(model_rfn200, test_data, type = "response")
df = confusionMatrix(predict_rfn, test_data[["Class"]])
df

precision <- df$byClass['Precision']    
precision
recall <- df$byClass['Recall']
recall
specificity<- df$byClass['Specificity']
specificity
F_Score = (2 * precision * recall) / (precision + recall)
paste("F-Score",F_Score)

```



```{r}
# Random Forest with SMOTE
model_rfn400 <-  randomForest(Class~., data = credit_smote,
                           ntree = 400, nodesize = 20)
#testing sample trained model on test data
predict_rfn <- predict(model_rfn400, test_data, type = "response")
df = confusionMatrix(predict_rfn, test_data[["Class"]])
df

precision <- df$byClass['Precision']    
precision
recall <- df$byClass['Recall']
recall
specificity<- df$byClass['Specificity']
specificity
F_Score = (2 * precision * recall) / (precision + recall)
paste("F-Score",F_Score)


```




```{r}
# Random Forest with SMOTE
#model_rfn
#set.seed(1)
#credit_smote = credit_smote %>% sample_frac(0.5)
nrow(credit_smote)
model_rfn3 <-  randomForest(Class~V14+V4, data = credit_smote)
#testing sample trained model on test data
predict_rfn4 <- predict(model_rfn3, test_data, type = "response")
confusionMatrix(predict_rfn4, test_data[["Class"]])

```



```{r}
# Random Forest with SMOTE
#model_rfn
#set.seed(1)
#credit_smote = credit_smote %>% sample_frac(0.5)
nrow(credit_smote)
model_rfn4 <-  randomForest(Class~V14+V4+V10+V12+V11, ntree=200, data = credit_smote)
#testing sample trained model on test data
predict_rfn5 <- predict(model_rfn4, test_data, type = "response")
df = confusionMatrix(predict_rfn5, test_data[["Class"]])
df
precision <- df$byClass['Precision']    
precision
recall <- df$byClass['Recall']
recall
specificity<- df$byClass['Specificity']
specificity
F_Score = (2 * precision * recall) / (precision + recall)
paste("F-Score",F_Score)

```

```{r}


#Predicting on Whole Test Dataset by taking 5 important features

predict_rfn6 <- predict(model_rfn4, credit_card[-1], type = "response")
df = confusionMatrix(predict_rfn6, credit_card[["Class"]])
df$table
df
precision <- df$byClass['Precision']    
precision

recall <- df$byClass['Recall']
recall
```

